---
output:
  bookdown::pdf_document2:
    toc: false
    number_sections: false
documentclass: article
geometry: margin=0.95in
classoption:
- onecolumn
bibliography: mama-mia.bib
csl: alex_vancouver.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, include=F)

library(tidyverse)
library(dplyr)
library(recipes)
library(kableExtra)
library(gtsummary)
library(ggplot2)
library(pROC)
library(precrec)
library(mice)

clinicalData <- read.csv("clinical_and_imaging_info.csv") %>% 
  mutate(
    pcr = replace_na(pcr, replace=0)
  )
relevantData <- clinicalData %>% 
  select(
    pcr, patient_id,
    dataset, nac_agent,
    bilateral_breast_cancer,
    multifocal_cancer,
    hr, er, pr, her2,
    nottingham_grade,
    tumor_subtype,
    age, weight, bmi_group,
    menopause, ethnicity
  )
```

```{r clean-data, cache=T}
# clean missing data
relevantData <- relevantData %>%
  mutate(across(where(is.character), ~na_if(str_trim(.), "")))

# clean up data
relevantData <- relevantData %>% select(
  patient_id, pcr, dataset, nac_agent,
  bilateral_breast_cancer,
  multifocal_cancer,
  her2, tumor_subtype,
  age, bmi_group,
  menopause, ethnicity
) %>% mutate(
  menopause = case_when(
    str_detect(menopause, "(?i)pre|peri") ~ "pre",
    str_detect(menopause, "(?i)post")     ~ "post",
    is.na(menopause)                      ~ NA_character_,
  ),
  tumor_subtype = case_when(
    str_detect(tumor_subtype, "(?i)luminal") ~ "luminal",
    str_detect(tumor_subtype, "(?i)her2")    ~ "her2_enriched",
    is.na(tumor_subtype)                     ~ NA_character_,
    TRUE                                     ~ tumor_subtype
  ),
  
  ethnicity = case_when(
    str_detect(ethnicity, "(?i)american indian|native") ~ "american indian/alaska native",
    str_detect(ethnicity, "(?i)hawai|pacific islander") ~ "hawaiian/pacific islander",
    str_detect(ethnicity, "(?i)multiple race")          ~ NA_character_,
    is.na(ethnicity)                                    ~ NA_character_,
    TRUE                                                ~ ethnicity
  )
)

# Now one-hot encode treatment variables
unique_agents <- relevantData %>%
  filter(!is.na(nac_agent) & nac_agent != "") %>%
  pull(nac_agent) %>%
  str_split(" \\+ ") %>%  # Splits by " + "
  unlist() %>%
  str_trim() %>%
  unique()

for (agent in unique_agents) {
  # Create a clean column name
  col_name <- paste0("agent_", str_replace_all(tolower(agent), "[^a-z0-9]", "_"))
  
  relevantData <- relevantData %>%
    mutate(!!col_name := case_when(
      is.na(nac_agent) ~ NA_real_,
      stringr::str_detect(as.character(nac_agent), stringr::fixed(agent)) ~ 1,
      TRUE ~ 0
    ))
}

relevantData <- relevantData %>% select(-nac_agent)
```

```{r cache=T}
write.csv(relevantData, "./processedData.csv", row.names=F)
```


<!-- Title here: -->

# \textit{Predicting Pathologic Complete Response from Clinical Variables}

<!-- Name here: -->

Tom Stoughton, MS Candidate[^1]

[^1]: Department of Biostatistics and Informatics, Colorado School of Public Health, University of Colorado Anschutz Medical Campus, Aurora, CO

## Abstract

Pathologic Complete Response (pCR) is a critical surrogate endpoint in neoadjuvant breast cancer treatment, traditionally used to evaluate the efficacy of systemic therapies. This study utilizes clinical, demographic, and treatment data from 1,506 patients to develop and compare predictive logistic regression models, specifically examining the trade-off between statistical performance and clinical utility. Using Multivariate Imputation by Chained Equations (MICE) to address missingness, we demonstrate that while the inclusion of specific treatment agents in Model 3 provides superior model parsimony and a slight increase in precision, baseline models focused on molecular subtypes (HER2 and Triple Negative) remain the most clinically significant. Model selection based on AIC and Precision-Recall curves indicates that treatment-augmented models refine predictions only after a therapeutic path has been chosen. In contrast, baseline models that exclude treatment variables are of higher clinical importance as they can be utilized at the point of diagnosis to inform the fundamental decision of whether neoadjuvant chemotherapy (NAC) is required at all. By identifying patients with a low probability of pCR regardless of agent, these models provide a framework for prioritizing surgical intervention and avoiding the debilitating side effects of chemotherapy when benefit is statistically minimal.

*Keywords: Pathologic Complete Response, Breast Cancer, Logistic Regression, *

## Introduction

Neoadjuvant chemotherapy (NAC) is standard of care for many breast cancer patients, aiming to shrink tumors prior to surgery. The achievement of Pathologic Complete Response (pCR)—defined as the absence of invasive cancer in the breast and lymph nodes at the time of surgery—is highly correlated with long-term survival. However, response rates vary significantly based on tumor biology and treatment regimen.
Predicting which patients will achieve pCR remains a challenge due to the complex interplay between patient demographics, tumor molecular profiles, and varied chemotherapy agents. This project aims to determine the incremental predictive value of adding demographic and treatment-specific data to a baseline clinical model.

## Dataset

This data set contains publicly available clinical and demographic information about 1,506 breast cancer patients sourced from four different clinical trials[@garrucho2025]. This dataset includes a robust array of diagnostic indicators essential for assessing treatment efficacy, such as HER2 status, Nottingham grade, and tumor subtype classifications including Triple Negative, Luminal, and HER2-enriched profiles. Clinical assessments also accounted for the presence of multifocal or bilateral disease to capture disease extent. Patient-level demographic variables integrated into the analysis include age, ethnicity, weight, BMI category, and menopausal status. For a significant subset of 1,100 patients, granular neoadjuvant chemotherapy (NAC) data were available, identifying the administration of specific therapeutic agents such as Pembrolizumab, Paclitaxel, Carboplatin, and Trastuzumab. 

To maintain the integrity and statistical power of the 1,506-patient sample, Multivariate Imputation by Chained Equations (MICE) (m=10 imputations) was utilized to address missingness, which was most prevalent in the menopause (~33%) and ethnicity variables. A Random Forest method was employed to capture non-linear relationships. Predictors included tumor subtype, HER2 status, BMI, age, and ethnicity.

## Methods

We employed a logistic regression framework to model the probability of pCR. The general form of the
model is:
logit(P(pCR = 1)) = β0
 + β1X1
 + β2X2
 + ... + βkXk
Where logit(p) = ln(p / (1-p)). We evaluated three distinct model parameters:
Model 1 (Baseline): Clinical features (Subtype, HER2, BMI, Menopause, Multifocality).
Model 2 (Demographic): Clinical features + Age and Ethnicity.
Model 3 (Treatment): All previous features + specific Chemotherapy Agents (e.g.,
Pembrolizumab, Carboplatin).

```{r read-processed-data}
data <- read.csv("./processedData.csv")
dataNoMissing <- data %>% drop_na()

```

```{r mice, cache=T, results='hide', message=F, warning=F}
# Impute the missing data (m=5 creates 5 versions of the data)
data_for_imputation <- data %>%
  mutate(across(where(is.character), as.factor)) %>%
  select(-starts_with("agent_"))

# 2. Initialize the setup without running the iterations yet
ini <- mice(data_for_imputation, maxit = 0)

# 3. Modify the Predictor Matrix
# Setting a column to 0 in the predictorMatrix means it won't be used to predict anything else
pred <- ini$predictorMatrix
pred[, "patient_id"] <- 0  # patient_id will not be used as a predictor
pred["patient_id", ] <- 0  # patient_id itself will not be imputed

# 4. Modify the Method
# Setting the method to "" tells mice to ignore this column during the math
meth <- ini$method
meth["patient_id"] <- ""

# 5. Run the actual imputation with these settings
imputed_data <- mice(data_for_imputation, 
                     m = 10, 
                     method = meth, 
                     predictorMatrix = pred, 
                     seed = 123)

```


```{r logistic-regression-clinical}
# Run your logistic regression on all 5 datasets and pool the results
fit1 <- with(imputed_data, 
             glm(pcr ~ multifocal_cancer + bilateral_breast_cancer + 
                 bmi_group + her2 + tumor_subtype + menopause, 
                 family = binomial))
pooled_results1 <- pool(fit1)

```

```{r clinical-results, include=T}
reg_table_data <- summary(pooled_results1) %>%
  mutate(across(where(is.numeric), ~ round(.x, 3))) %>%
  rename(Variable = term, P_Value = p.value)

# 2. Identify rows where p-value < 0.05
sig_rows <- which(reg_table_data$P_Value < 0.05)

# 3. Create the table
reg_table_data %>%
  kbl(caption = "Model 1: Pooled Logistic Regression",
      booktabs = TRUE) %>%
  kable_styling(bootstrap_options = c("striped", "hover"), 
                full_width = FALSE, 
                latex_options = "HOLD_position") %>%
  row_spec(sig_rows, bold = TRUE) %>% # Bolds significant predictors like Pembrolizumab
  column_spec(1, italic = TRUE)

```

```{r logistic-regression-demographic}
# Run your logistic regression on all 5 datasets and pool the results
fit2 <- with(imputed_data, 
             glm(pcr ~ multifocal_cancer + bilateral_breast_cancer + 
                  bmi_group + age + her2 + tumor_subtype + 
                  menopause + ethnicity, family = binomial))
pooled_results2 <- pool(fit2)

```

```{r demographic-results, include=T}
reg_table_data <- summary(pooled_results2) %>%
  mutate(across(where(is.numeric), ~ round(.x, 3))) %>%
  rename(Variable = term, P_Value = p.value)

# 2. Identify rows where p-value < 0.05
sig_rows <- which(reg_table_data$P_Value < 0.05)

# 3. Create the table
reg_table_data %>%
  kbl(caption = "Model 2: Pooled Logistic Regression (with Demographics)",
      booktabs = TRUE) %>%
  kable_styling(bootstrap_options = c("striped", "hover"), 
                full_width = FALSE, 
                latex_options = "HOLD_position") %>%
  row_spec(sig_rows, bold = TRUE) %>% # Bolds significant predictors like Pembrolizumab
  column_spec(1, italic = TRUE)

```

```{r logistic-regression-agent}
# 1. Identify agent columns
agents_to_join <- data %>% 
  mutate(.id = as.integer(row_number())) %>% 
  select(.id, starts_with("agent_"))

subset_long_data <- complete(imputed_data, action = "long", include = TRUE) %>%
  left_join(agents_to_join, by = ".id") %>%
  filter(dataset %in% c("ISPY2", "NACT")) %>%
  mutate(ethnicity = as.character(ethnicity),
         # Group Hispanic into 'Other' or 'Caucasian' (or whatever your baseline is)
         ethnicity = ifelse(ethnicity == "hispanic", "Other/Unknown", ethnicity),
         ethnicity = as.factor(ethnicity)) %>%
  select(pcr, .imp, .id, everything()) %>% # Bring pcr to the front to be sure
  mutate(.imp = as.integer(.imp), # as.mids often prefers integer .imp for indexing
         .id = as.integer(.id))

agent_cols <- names(subset_long_data)[grep("agent_", names(subset_long_data))]

# 2. Calculate Correlation Matrix
cor_matrix <- cor(subset_long_data[, agent_cols], use = "pairwise.complete.obs")

processed <- c()
collapsed_data <- subset_long_data

for (col in agent_cols) {
  if (col %in% processed) next
  
  # A. Identify Positive Collinearity (r > 0.99)
  identical_cols <- names(which(cor_matrix[col, ] > 0.99))
  
  # B. Identify Anti-Collinearity (r < -0.99)
  anti_cols <- names(which(cor_matrix[col, ] < -0.99))
  
  if (length(identical_cols) > 1) {
    # Combine Positive Collinearity
    new_name <- paste(identical_cols[1], 
                      paste(gsub("agent_", "", identical_cols[-1]), collapse = "_"), 
                      sep = "_")
    collapsed_data <- collapsed_data %>%
      mutate(!!new_name := get(col)) %>%
      select(-all_of(identical_cols))
    processed <- c(processed, identical_cols)
    
  } else if (length(anti_cols) > 0) {
    # Combine Anti-Collinearity (Mutually Exclusive)
    # Name it: agent_DrugA_vs_DrugB
    new_name <- paste0(col, "_vs_", gsub("agent_", "", anti_cols[1]))
    
    collapsed_data <- collapsed_data %>%
      mutate(!!new_name := get(col)) %>% # 1 = DrugA, 0 = DrugB
      select(-all_of(c(col, anti_cols)))
    
    processed <- c(processed, col, anti_cols)
  }
}

# 3. Re-run MICE object creation
imputed_collapsed <- as.mids(collapsed_data)

# 4. Update your formula and fit
new_agent_vars <- names(collapsed_data)[grep("agent_", names(collapsed_data))]
final_agent_vars <- new_agent_vars[!new_agent_vars %in% c("agent_fec100")]
formula_str <- paste("pcr ~ bilateral_breast_cancer + multifocal_cancer + her2 + 
                     tumor_subtype + age + bmi_group + menopause + ethnicity +", 
                     paste(final_agent_vars, collapse = " + "))

fit3 <- with(imputed_collapsed, glm(as.formula(formula_str), family = binomial))
pooled_results3 <- pool(fit3)

```


```{r agent-results, include=T}
reg_table_data <- summary(pooled_results3) %>%
  mutate(across(where(is.numeric), ~ round(.x, 3))) %>%
  rename(Variable = term, P_Value = p.value)

# 2. Identify rows where p-value < 0.05
sig_rows <- which(reg_table_data$P_Value < 0.05)

# 3. Create the table
reg_table_data %>%
  kbl(caption = "Model 3: Pooled Logistic Regression (Clinical + Treatment Agents)",
      booktabs = TRUE) %>%
  kable_styling(bootstrap_options = c("striped", "hover"), 
                full_width = FALSE, 
                latex_options = "HOLD_position") %>%
  row_spec(sig_rows, bold = TRUE) %>% # Bolds significant predictors like Pembrolizumab
  column_spec(1, italic = TRUE)

```

## Results

```{r auroc-average-1, results=F, include=T, message=F, width=8, height=8}
# 1. Get predicted probabilities from each of the 5 models
# This creates a matrix of probabilities (one column per imputation)
predictions <- list()
for(i in 1:imputed_data$m) {
  # Extract the i-th completed dataset
  dat <- complete(imputed_data, i)
  
  # Fit the model on that specific dataset
  m_fit <- glm(pcr ~ multifocal_cancer + bilateral_breast_cancer + 
                 bmi_group + her2 + tumor_subtype + menopause, 
               data = dat, family = binomial)
  
  # Store the predicted probabilities
  predictions[[i]] <- predict(m_fit, type = "response")
}

# 2. Average the probabilities across the 5 imputations
avg_probs_m1 <- rowMeans(as.data.frame(predictions))

# 3. Calculate AUC using the actual 'pcr' values from any completed dataset 
# (pcr is the same in all of them since it wasn't imputed)
actual_pcr <- complete(imputed_data, 1)$pcr
roc_curve <- roc(actual_pcr, avg_probs_m1)
m1_auc <- pROC::auc(roc_curve)

# 4. Results
par(pty = "s") 

# 2. Plot the ROC curve with your specific requirements
plot(roc_curve, 
     col = "brown1", 
     lwd = 3,                       # Make the line thicker for readability
     main = paste("Pooled AUC-ROC for Model 1:", round(m1_auc, 3)),
     
     # Axis Labels
     xlab = "1-FPR",                # Your requested X-axis label
     ylab = "TPR",                  # Your requested Y-axis label
     
     # Bounds and Scaling
     xlim = c(1, 0),                # Standard ROC X-axis is 1 to 0 (FPR decreases)
     ylim = c(0, 1),                # Y-axis bounds
     
     # Formatting
     legacy.axes = TRUE,            # Ensures the X-axis goes from 0 to 1 if preferred
     asp = 1,                       # Ensures an aspect ratio of 1 (strictly square)
     grid = TRUE                    # Adds a grid for easier value reading
)

# 3. Reset the plot parameters to default for future plots
par(pty = "m")

```

```{r auroc-average-2, results=F, include=T, message=F, width=8, height=8}
# 1. Get predicted probabilities from each of the 5 models
# This creates a matrix of probabilities (one column per imputation)
predictions <- list()
for(i in 1:imputed_data$m) {
  # Extract the i-th completed dataset
  dat <- complete(imputed_data, i)
  
  # Fit the model on that specific dataset
  m_fit <- glm(pcr ~ multifocal_cancer + bilateral_breast_cancer + 
                  bmi_group + age + her2 + tumor_subtype + 
                  menopause + ethnicity, 
               data = dat, family = binomial)
  
  # Store the predicted probabilities
  predictions[[i]] <- predict(m_fit, type = "response")
}

# 2. Average the probabilities across the 5 imputations
avg_probs_m2 <- rowMeans(as.data.frame(predictions))

# 3. Calculate AUC using the actual 'pcr' values from any completed dataset 
# (pcr is the same in all of them since it wasn't imputed)
actual_pcr <- complete(imputed_data, 1)$pcr
roc_curve <- roc(actual_pcr, avg_probs_m2)
m2_auc <- pROC::auc(roc_curve)

# 4. Results
par(pty = "s") 

plot(roc_curve, 
     col = "royalblue", 
     lwd = 3,                       # Make the line thicker for readability
     main = paste("Pooled AUC-ROC for Model 2:", round(m2_auc, 3)),
     
     # Axis Labels
     xlab = "1-FPR",                # Your requested X-axis label
     ylab = "TPR",                  # Your requested Y-axis label
     
     # Bounds and Scaling
     xlim = c(1, 0),                # Standard ROC X-axis is 1 to 0 (FPR decreases)
     ylim = c(0, 1),                # Y-axis bounds
     
     # Formatting
     legacy.axes = TRUE,            # Ensures the X-axis goes from 0 to 1 if preferred
     asp = 1,                       # Ensures an aspect ratio of 1 (strictly square)
     grid = TRUE                    # Adds a grid for easier value reading
)

par(pty = "m")

```

```{r auroc-average-3, results=F, include=T, message=F, width=8, height=8}
predictions <- list()

for(i in 1:imputed_collapsed$m) {
  # Extract the i-th completed dataset (which already has agents and filtering)
  dat <- complete(imputed_collapsed, i)
  
  # Fit the treatment model (Model 3)
  # We use the explicit variable list to ensure it matches your pooled summary
  m_fit <- glm(pcr ~ bilateral_breast_cancer + multifocal_cancer + 
                 her2 + tumor_subtype + age + bmi_group + 
                 menopause + ethnicity +
                 agent_paclitaxel + agent_ganitumab + agent_ganetespib + 
                 agent_mk_2206 + agent_pembrolizumab + agent_amg_386 + 
                 agent_trastuzumab + agent_neratinib + agent_abt_888_carboplatin + 
                 agent_pertuzumab + agent_taxane, 
               data = dat, family = binomial)
  
  # Store the predicted probabilities
  predictions[[i]] <- predict(m_fit, type = "response")
}

# 2. Average the probabilities across the 5 imputations
avg_probs_m3 <- rowMeans(as.data.frame(predictions))

# 3. Calculate AUC using the actual 'pcr' values from the subset
# We pull pcr from the first imputation of the SUBSET
actual_pcr <- complete(imputed_collapsed, 1)$pcr
roc_curve_m3 <- roc(actual_pcr, avg_probs_m3)
m3_auc <- pROC::auc(roc_curve_m3)

# 4. Results
par(pty = "s") 

plot(roc_curve_m3, 
     col = "darkgreen", 
     lwd = 3,                       # Make the line thicker for readability
     main = paste("Pooled AUC-ROC for Model 3:", round(m3_auc, 3)),
     
     # Axis Labels
     xlab = "1-FPR",                # Your requested X-axis label
     ylab = "TPR",                  # Your requested Y-axis label
     
     # Bounds and Scaling
     xlim = c(1, 0),                # Standard ROC X-axis is 1 to 0 (FPR decreases)
     ylim = c(0, 1),                # Y-axis bounds
     
     # Formatting
     legacy.axes = TRUE,            # Ensures the X-axis goes from 0 to 1 if preferred
     asp = 1,                       # Ensures an aspect ratio of 1 (strictly square)
     grid = TRUE                    # Adds a grid for easier value reading
)

# 3. Reset the plot parameters to default for future plots
par(pty = "m")

```

```{r aic-plots}
# --- STAGE 1: Identify the clinical Subset IDs ---

# Extract the unique patient IDs from the ISPY2/NACT treatment subset
# We pull from the first imputation since patient_id is consistent across all
target_patient_ids <- complete(imputed_collapsed, 1)$patient_id

# --- STAGE 2: Refit Models 1 & 2 using the patient_id filter ---

# We refit on the full imputed_data object, but restrict it to the treatment IDs.
# Note: 'with' automatically handles the looping through all 10 imputations.
fit1_refit <- with(imputed_data, 
                   glm(pcr ~ multifocal_cancer + bilateral_breast_cancer + 
                       bmi_group + her2 + tumor_subtype + menopause, 
                       subset = patient_id %in% target_patient_ids, 
                       family = binomial))

fit2_refit <- with(imputed_data, 
                   glm(pcr ~ multifocal_cancer + bilateral_breast_cancer + 
                       bmi_group + her2 + tumor_subtype + menopause + age + 
                       ethnicity, 
                       subset = patient_id %in% target_patient_ids, 
                       family = binomial))

# --- STAGE 3: Calculate the Valid AIC Comparison ---

# We compare these refitted models to your Model 3 (fit3)
aic_comparison <- data.frame(
  Model = c("Model 1 (Baseline)", "Model 2 (Demographic)", "Model 3 (Treatment)"),
  AIC = c(
    mean(sapply(fit1_refit$analyses, AIC)),
    mean(sapply(fit2_refit$analyses, AIC)),
    mean(sapply(fit3$analyses, AIC))
  )
)

# --- STAGE 4: Plot the Standardized Comparison ---
ggplot(aic_comparison, aes(x = Model, y = AIC, fill = Model)) +
  geom_bar(stat = "identity", width = 0.5, show.legend = FALSE) +
  geom_text(aes(label = round(AIC, 1)), vjust = -0.5, fontface = "bold") +
  theme_minimal() +
  labs(title = "Model Selection: Standardized AIC Comparison",
       subtitle = paste0("N = ", length(target_patient_ids), " patients (ISPY2/NACT Subset)"),
       y = "Mean AIC",
       x = "") +
  scale_color_manual(values = c("brown1", "royalblue", "darkgreen")) +
  scale_fill_brewer(palette = "Set1")
```

```{r prec-recall, warning=F}
# 1. Helper function to safely extract averaged probabilities from a 'mira' object
get_avg_probs <- function(mira_fit) {
  # Get the number of imputations (m)
  m_count <- length(mira_fit$analyses)
  
  # Extract predictions from each of the m models
  pred_list <- lapply(mira_fit$analyses, function(x) predict(x, type = "response"))
  
  # Average them across rows
  rowMeans(as.data.frame(pred_list))
}

# 2. Calculate probabilities for the two baseline models (Refit on Subset)
avg_probs_m1 <- get_avg_probs(fit1_refit)
avg_probs_m2 <- get_avg_probs(fit2_refit)

# 3. Get the actual pCR values from the subset
actual_pcr <- complete(imputed_collapsed, 1)$pcr

# 4. Create the PR multi-plot object
precrec_obj <- evalmod(
  scores = list(avg_probs_m1, avg_probs_m2, avg_probs_m3), 
  labels = list(actual_pcr, actual_pcr, actual_pcr),
  modnames = c("Model 1: Baseline", "Model 2: Demographic", "Model 3: Treatment")
)

# 5. Plot the Comparison
autoplot(precrec_obj, curvetype = "PRC") +
  theme_minimal() +
  scale_color_manual(values = c("brown1", "royalblue", "darkgreen")) +
  labs(title = "Precision-Recall Comparison: Clinical Impact",
       subtitle = "Evaluation of added predictive value on standardized subset",
       x = "Recall (Sensitivity)",
       y = "Precision (PPV)")
```

```{r forest-plot, warning=F}
# 1. Get the pooled results into a standard dataframe
pooled_df <- as.data.frame(summary(pool(fit3), conf.int = TRUE))

# 2. Rename columns manually if they aren't standard
# Depending on your mice version, column names might be 'estimate' or 'est'
names(pooled_df)[names(pooled_df) == "estimate"] <- "est"
names(pooled_df)[names(pooled_df) == "2.5 %"] <- "low"
names(pooled_df)[names(pooled_df) == "97.5 %"] <- "high"

# 3. Exclude intercept and transform to Odds Ratios
plot_data <- pooled_df[pooled_df$term != "(Intercept)", ]
plot_data$OR <- exp(plot_data$est)
plot_data$OR_low <- exp(plot_data$low)
plot_data$OR_high <- exp(plot_data$high)

# 4. Generate the Forest Plot using standard ggplot2
ggplot(plot_data, aes(x = reorder(term, OR), y = OR)) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "steelblue") +
  geom_pointrange(aes(ymin = OR_low, ymax = OR_high), color = "darkgreen") +
  coord_flip() +
  theme_bw() + # theme_bw is very stable in older ggplot2 versions
  labs(title = "Model 3: Odds Ratios for pCR",
       x = "Predictor", 
       y = "Odds Ratio (95% CI)") +
  scale_y_log10()

```


## Conclusion

The comparative analysis of these three models reveals a critical distinction between statistical optimization and clinical utility. While the inclusion of neoadjuvant treatment agents in Model 3 provides the highest predictive precision and the most parsimonious fit according to AIC metrics, this performance gain must be weighed against practical application in a pre-treatment setting. The treatment-augmented model is inherently retrospective; it confirms that specific drugs drive the pathologic response, effectively mediating the biological signals of subtypes like Triple Negative.  
However, for the practicing clinician, the baseline and demographic models (Models 1 and 2) hold greater clinical importance. By relying solely on information available before a treatment plan is finalized, these models provide the necessary foresight to evaluate whether a patient is likely to benefit from neoadjuvant chemotherapy at all. In cases where baseline models predict a negligible probability of pCR, clinicians may consider bypassing systemic therapy in favor of immediate surgical intervention. This approach prioritizes patient quality of life by avoiding the significant toxicities and long-term side effects of chemotherapy when the likelihood of a complete pathologic response is statistically minimal. Ultimately, while treatment variables slightly increase the predictive "ceiling," the true value of predictive modeling in this cohort lies in its ability to support the early de-escalation of care. 

\newpage
\onecolumn

## References

\vspace{2mm}

::: {#refs}
:::

## Code Appendix

```{r, ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE}
```
