---
output:
  bookdown::pdf_document2:
    toc: false
    number_sections: false
documentclass: article
geometry: margin=0.95in
classoption:
- onecolumn
bibliography: mama-mia.bib
csl: alex_vancouver.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)

library(tidyverse)
library(dplyr)
library(recipes)
library(kableExtra)
library(gtsummary)
library(ggplot2)
library(pROC)
library(precrec)
library(mice)

clinicalData <- read.csv("clinical_and_imaging_info.csv") %>% 
  mutate(
    pcr = replace_na(pcr, replace=0)
  )
relevantData <- clinicalData %>% 
  select(
    pcr, patient_id,
    dataset, nac_agent,
    bilateral_breast_cancer,
    multifocal_cancer,
    hr, er, pr, her2,
    nottingham_grade,
    tumor_subtype,
    age, weight, bmi_group,
    menopause, ethnicity
  )
```

```{r clean-data, cache=T}
# clean missing data
relevantData <- relevantData %>%
  mutate(across(where(is.character), ~na_if(str_trim(.), "")))

# clean up data
relevantData <- relevantData %>% select(
  patient_id, pcr, dataset, nac_agent,
  bilateral_breast_cancer,
  multifocal_cancer,
  her2, tumor_subtype,
  age, bmi_group,
  menopause, ethnicity
) %>% mutate(
  menopause = case_when(
    str_detect(menopause, "(?i)pre|peri") ~ "pre",
    str_detect(menopause, "(?i)post")     ~ "post",
    is.na(menopause)                      ~ NA_character_,
  ),
  tumor_subtype = case_when(
    str_detect(tumor_subtype, "(?i)luminal") ~ "luminal",
    str_detect(tumor_subtype, "(?i)her2")    ~ "her2_enriched",
    is.na(tumor_subtype)                     ~ NA_character_,
    TRUE                                     ~ tumor_subtype
  ),
  
  ethnicity = case_when(
    str_detect(ethnicity, "(?i)american indian|native") ~ "american indian/alaska native",
    str_detect(ethnicity, "(?i)hawai|pacific islander") ~ "hawaiian/pacific islander",
    str_detect(ethnicity, "(?i)multiple race")          ~ NA_character_,
    is.na(ethnicity)                                    ~ NA_character_,
    TRUE                                                ~ ethnicity
  )
)

# Now one-hot encode treatment variables
unique_agents <- relevantData %>%
  filter(!is.na(nac_agent) & nac_agent != "") %>%
  pull(nac_agent) %>%
  str_split(" \\+ ") %>%  # Splits by " + "
  unlist() %>%
  str_trim() %>%
  unique()

for (agent in unique_agents) {
  # Create a clean column name
  col_name <- paste0("agent_", str_replace_all(tolower(agent), "[^a-z0-9]", "_"))
  
  relevantData <- relevantData %>%
    mutate(!!col_name := case_when(
      is.na(nac_agent) ~ NA_real_,
      stringr::str_detect(as.character(nac_agent), stringr::fixed(agent)) ~ 1,
      TRUE ~ 0
    ))
}

relevantData <- relevantData %>% select(-nac_agent)
```

```{r cache=T}
write.csv(relevantData, "./processedData.csv", row.names=F)
```


<!-- Title here: -->

# \textit{Predicting Pathologic Complete Response from Clinical Variables}

<!-- Name here: -->

Tom Stoughton, MS Candidate[^1]

[^1]: Department of Biostatistics and Informatics, Colorado School of Public Health, University of Colorado Anschutz Medical Campus, Aurora, CO

## Abstract

...

*Keywords: Pathologic Complete Response, Breast Cancer, Logistic Regression, *

## Introduction

This data set contains publicly available clinical and demographic information about 1,506 breast cancer patients from four different trials[@garrucho2025]

## Motivating Problem

...

## Methods

```{r read-processed-data}
data <- read.csv("./processedData.csv")
dataNoMissing <- data %>% drop_na()

```

Using Multivariate Imputation by Chained Equations (MICE).

```{r mice, cache=T, results='hide', message=F, warning=F}
# Impute the missing data (m=5 creates 5 versions of the data)
data_for_imputation <- data %>%
  mutate(across(where(is.character), as.factor)) %>%
  select(-starts_with("agent_"))

# 2. Initialize the setup without running the iterations yet
ini <- mice(data_for_imputation, maxit = 0)

# 3. Modify the Predictor Matrix
# Setting a column to 0 in the predictorMatrix means it won't be used to predict anything else
pred <- ini$predictorMatrix
pred[, "patient_id"] <- 0  # patient_id will not be used as a predictor
pred["patient_id", ] <- 0  # patient_id itself will not be imputed

# 4. Modify the Method
# Setting the method to "" tells mice to ignore this column during the math
meth <- ini$method
meth["patient_id"] <- ""

# 5. Run the actual imputation with these settings
imputed_data <- mice(data_for_imputation, 
                     m = 10, 
                     method = meth, 
                     predictorMatrix = pred, 
                     seed = 123)

```

logistic regression on clinical variables.

```{r logistic-regression-clinical}
# Run your logistic regression on all 5 datasets and pool the results
fit1 <- with(imputed_data, 
             glm(pcr ~ multifocal_cancer + bilateral_breast_cancer + 
                 bmi_group + her2 + tumor_subtype + menopause, 
                 family = binomial))
pooled_results1 <- pool(fit1)

```

```{r clinical-results, include=T}
reg_table_data <- summary(pooled_results1) %>%
  mutate(across(where(is.numeric), ~ round(.x, 3))) %>%
  rename(Variable = term, P_Value = p.value)

# 2. Identify rows where p-value < 0.05
sig_rows <- which(reg_table_data$P_Value < 0.05)

# 3. Create the table
reg_table_data %>%
  kbl(caption = "Model 1: Pooled Logistic Regression",
      booktabs = TRUE) %>%
  kable_styling(bootstrap_options = c("striped", "hover"), 
                full_width = FALSE, 
                latex_options = "HOLD_position") %>%
  row_spec(sig_rows, bold = TRUE) %>% # Bolds significant predictors like Pembrolizumab
  column_spec(1, italic = TRUE)

```

Now for including demographics (age and ethnicity).

```{r logistic-regression-demographic}
# Run your logistic regression on all 5 datasets and pool the results
fit2 <- with(imputed_data, 
             glm(pcr ~ multifocal_cancer + bilateral_breast_cancer + 
                  bmi_group + age + her2 + tumor_subtype + 
                  menopause + ethnicity, family = binomial))
pooled_results2 <- pool(fit2)

```

```{r demographic-results, include=T}
reg_table_data <- summary(pooled_results2) %>%
  mutate(across(where(is.numeric), ~ round(.x, 3))) %>%
  rename(Variable = term, P_Value = p.value)

# 2. Identify rows where p-value < 0.05
sig_rows <- which(reg_table_data$P_Value < 0.05)

# 3. Create the table
reg_table_data %>%
  kbl(caption = "Model 2: Pooled Logistic Regression (with Demographics)",
      booktabs = TRUE) %>%
  kable_styling(bootstrap_options = c("striped", "hover"), 
                full_width = FALSE, 
                latex_options = "HOLD_position") %>%
  row_spec(sig_rows, bold = TRUE) %>% # Bolds significant predictors like Pembrolizumab
  column_spec(1, italic = TRUE)

```

Now for including treatment agents.

```{r logistic-regression-agent}
# 1. Identify agent columns
agents_to_join <- data %>% 
  mutate(.id = as.integer(row_number())) %>% 
  select(.id, starts_with("agent_"))

subset_long_data <- complete(imputed_data, action = "long", include = TRUE) %>%
  left_join(agents_to_join, by = ".id") %>%
  filter(dataset %in% c("ISPY2", "NACT")) %>%
  mutate(ethnicity = as.character(ethnicity),
         # Group Hispanic into 'Other' or 'Caucasian' (or whatever your baseline is)
         ethnicity = ifelse(ethnicity == "hispanic", "Other/Unknown", ethnicity),
         ethnicity = as.factor(ethnicity)) %>%
  select(pcr, .imp, .id, everything()) %>% # Bring pcr to the front to be sure
  mutate(.imp = as.integer(.imp), # as.mids often prefers integer .imp for indexing
         .id = as.integer(.id))

agent_cols <- names(subset_long_data)[grep("agent_", names(subset_long_data))]

# 2. Calculate Correlation Matrix
cor_matrix <- cor(subset_long_data[, agent_cols], use = "pairwise.complete.obs")

processed <- c()
collapsed_data <- subset_long_data

for (col in agent_cols) {
  if (col %in% processed) next
  
  # A. Identify Positive Collinearity (r > 0.99)
  identical_cols <- names(which(cor_matrix[col, ] > 0.99))
  
  # B. Identify Anti-Collinearity (r < -0.99)
  anti_cols <- names(which(cor_matrix[col, ] < -0.99))
  
  if (length(identical_cols) > 1) {
    # Combine Positive Collinearity
    new_name <- paste(identical_cols[1], 
                      paste(gsub("agent_", "", identical_cols[-1]), collapse = "_"), 
                      sep = "_")
    collapsed_data <- collapsed_data %>%
      mutate(!!new_name := get(col)) %>%
      select(-all_of(identical_cols))
    processed <- c(processed, identical_cols)
    
  } else if (length(anti_cols) > 0) {
    # Combine Anti-Collinearity (Mutually Exclusive)
    # Name it: agent_DrugA_vs_DrugB
    new_name <- paste0(col, "_vs_", gsub("agent_", "", anti_cols[1]))
    
    collapsed_data <- collapsed_data %>%
      mutate(!!new_name := get(col)) %>% # 1 = DrugA, 0 = DrugB
      select(-all_of(c(col, anti_cols)))
    
    processed <- c(processed, col, anti_cols)
  }
}

# 3. Re-run MICE object creation
imputed_collapsed <- as.mids(collapsed_data)

# 4. Update your formula and fit
new_agent_vars <- names(collapsed_data)[grep("agent_", names(collapsed_data))]
final_agent_vars <- new_agent_vars[!new_agent_vars %in% c("agent_fec100")]
formula_str <- paste("pcr ~ bilateral_breast_cancer + multifocal_cancer + her2 + 
                     tumor_subtype + age + bmi_group + menopause + ethnicity +", 
                     paste(final_agent_vars, collapse = " + "))

fit3 <- with(imputed_collapsed, glm(as.formula(formula_str), family = binomial))
pooled_results3 <- pool(fit3)

```


```{r agent-results, include=T}
reg_table_data <- summary(pooled_results3) %>%
  mutate(across(where(is.numeric), ~ round(.x, 3))) %>%
  rename(Variable = term, P_Value = p.value)

# 2. Identify rows where p-value < 0.05
sig_rows <- which(reg_table_data$P_Value < 0.05)

# 3. Create the table
reg_table_data %>%
  kbl(caption = "Model 3: Pooled Logistic Regression (Clinical + Treatment Agents)",
      booktabs = TRUE) %>%
  kable_styling(bootstrap_options = c("striped", "hover"), 
                full_width = FALSE, 
                latex_options = "HOLD_position") %>%
  row_spec(sig_rows, bold = TRUE) %>% # Bolds significant predictors like Pembrolizumab
  column_spec(1, italic = TRUE)

```

## Results

```{r auroc-average-1, results=F, include=T, message=F, width=8, height=8}
# 1. Get predicted probabilities from each of the 5 models
# This creates a matrix of probabilities (one column per imputation)
predictions <- list()
for(i in 1:imputed_data$m) {
  # Extract the i-th completed dataset
  dat <- complete(imputed_data, i)
  
  # Fit the model on that specific dataset
  m_fit <- glm(pcr ~ multifocal_cancer + bilateral_breast_cancer + 
                 bmi_group + her2 + tumor_subtype + menopause, 
               data = dat, family = binomial)
  
  # Store the predicted probabilities
  predictions[[i]] <- predict(m_fit, type = "response")
}

# 2. Average the probabilities across the 5 imputations
avg_probs_m1 <- rowMeans(as.data.frame(predictions))

# 3. Calculate AUC using the actual 'pcr' values from any completed dataset 
# (pcr is the same in all of them since it wasn't imputed)
actual_pcr <- complete(imputed_data, 1)$pcr
roc_curve <- roc(actual_pcr, avg_probs_m1)
m1_auc <- pROC::auc(roc_curve)

# 4. Results
par(pty = "s") 

# 2. Plot the ROC curve with your specific requirements
plot(roc_curve, 
     col = "brown1", 
     lwd = 3,                       # Make the line thicker for readability
     main = paste("Pooled AUC-ROC for Model 1:", round(m1_auc, 3)),
     
     # Axis Labels
     xlab = "1-FPR",                # Your requested X-axis label
     ylab = "TPR",                  # Your requested Y-axis label
     
     # Bounds and Scaling
     xlim = c(1, 0),                # Standard ROC X-axis is 1 to 0 (FPR decreases)
     ylim = c(0, 1),                # Y-axis bounds
     
     # Formatting
     legacy.axes = TRUE,            # Ensures the X-axis goes from 0 to 1 if preferred
     asp = 1,                       # Ensures an aspect ratio of 1 (strictly square)
     grid = TRUE                    # Adds a grid for easier value reading
)

# 3. Reset the plot parameters to default for future plots
par(pty = "m")

```

```{r auroc-average-2, results=F, include=T, message=F, width=8, height=8}
# 1. Get predicted probabilities from each of the 5 models
# This creates a matrix of probabilities (one column per imputation)
predictions <- list()
for(i in 1:imputed_data$m) {
  # Extract the i-th completed dataset
  dat <- complete(imputed_data, i)
  
  # Fit the model on that specific dataset
  m_fit <- glm(pcr ~ multifocal_cancer + bilateral_breast_cancer + 
                  bmi_group + age + her2 + tumor_subtype + 
                  menopause + ethnicity, 
               data = dat, family = binomial)
  
  # Store the predicted probabilities
  predictions[[i]] <- predict(m_fit, type = "response")
}

# 2. Average the probabilities across the 5 imputations
avg_probs_m2 <- rowMeans(as.data.frame(predictions))

# 3. Calculate AUC using the actual 'pcr' values from any completed dataset 
# (pcr is the same in all of them since it wasn't imputed)
actual_pcr <- complete(imputed_data, 1)$pcr
roc_curve <- roc(actual_pcr, avg_probs_m2)
m2_auc <- pROC::auc(roc_curve)

# 4. Results
par(pty = "s") 

plot(roc_curve, 
     col = "royalblue", 
     lwd = 3,                       # Make the line thicker for readability
     main = paste("Pooled AUC-ROC for Model 2:", round(m2_auc, 3)),
     
     # Axis Labels
     xlab = "1-FPR",                # Your requested X-axis label
     ylab = "TPR",                  # Your requested Y-axis label
     
     # Bounds and Scaling
     xlim = c(1, 0),                # Standard ROC X-axis is 1 to 0 (FPR decreases)
     ylim = c(0, 1),                # Y-axis bounds
     
     # Formatting
     legacy.axes = TRUE,            # Ensures the X-axis goes from 0 to 1 if preferred
     asp = 1,                       # Ensures an aspect ratio of 1 (strictly square)
     grid = TRUE                    # Adds a grid for easier value reading
)

par(pty = "m")

```

```{r auroc-average-3, results=F, include=T, message=F, width=8, height=8}
predictions <- list()

for(i in 1:imputed_collapsed$m) {
  # Extract the i-th completed dataset (which already has agents and filtering)
  dat <- complete(imputed_collapsed, i)
  
  # Fit the treatment model (Model 3)
  # We use the explicit variable list to ensure it matches your pooled summary
  m_fit <- glm(pcr ~ bilateral_breast_cancer + multifocal_cancer + 
                 her2 + tumor_subtype + age + bmi_group + 
                 menopause + ethnicity +
                 agent_paclitaxel + agent_ganitumab + agent_ganetespib + 
                 agent_mk_2206 + agent_pembrolizumab + agent_amg_386 + 
                 agent_trastuzumab + agent_neratinib + agent_abt_888_carboplatin + 
                 agent_pertuzumab + agent_taxane, 
               data = dat, family = binomial)
  
  # Store the predicted probabilities
  predictions[[i]] <- predict(m_fit, type = "response")
}

# 2. Average the probabilities across the 5 imputations
avg_probs_m3 <- rowMeans(as.data.frame(predictions))

# 3. Calculate AUC using the actual 'pcr' values from the subset
# We pull pcr from the first imputation of the SUBSET
actual_pcr <- complete(imputed_collapsed, 1)$pcr
roc_curve_m3 <- roc(actual_pcr, avg_probs_m3)
m3_auc <- pROC::auc(roc_curve_m3)

# 4. Results
par(pty = "s") 

plot(roc_curve_m3, 
     col = "darkgreen", 
     lwd = 3,                       # Make the line thicker for readability
     main = paste("Pooled AUC-ROC for Model 3:", round(m3_auc, 3)),
     
     # Axis Labels
     xlab = "1-FPR",                # Your requested X-axis label
     ylab = "TPR",                  # Your requested Y-axis label
     
     # Bounds and Scaling
     xlim = c(1, 0),                # Standard ROC X-axis is 1 to 0 (FPR decreases)
     ylim = c(0, 1),                # Y-axis bounds
     
     # Formatting
     legacy.axes = TRUE,            # Ensures the X-axis goes from 0 to 1 if preferred
     asp = 1,                       # Ensures an aspect ratio of 1 (strictly square)
     grid = TRUE                    # Adds a grid for easier value reading
)

# 3. Reset the plot parameters to default for future plots
par(pty = "m")

```

```{r aic-plots}
# --- STAGE 1: Identify the clinical Subset IDs ---

# Extract the unique patient IDs from the ISPY2/NACT treatment subset
# We pull from the first imputation since patient_id is consistent across all
target_patient_ids <- complete(imputed_collapsed, 1)$patient_id

# --- STAGE 2: Refit Models 1 & 2 using the patient_id filter ---

# We refit on the full imputed_data object, but restrict it to the treatment IDs.
# Note: 'with' automatically handles the looping through all 10 imputations.
fit1_refit <- with(imputed_data, 
                   glm(pcr ~ multifocal_cancer + bilateral_breast_cancer + 
                       bmi_group + her2 + tumor_subtype + menopause, 
                       subset = patient_id %in% target_patient_ids, 
                       family = binomial))

fit2_refit <- with(imputed_data, 
                   glm(pcr ~ multifocal_cancer + bilateral_breast_cancer + 
                       bmi_group + her2 + tumor_subtype + menopause + age + 
                       ethnicity, 
                       subset = patient_id %in% target_patient_ids, 
                       family = binomial))

# --- STAGE 3: Calculate the Valid AIC Comparison ---

# We compare these refitted models to your Model 3 (fit3)
aic_comparison <- data.frame(
  Model = c("Model 1 (Baseline)", "Model 2 (Demographic)", "Model 3 (Treatment)"),
  AIC = c(
    mean(sapply(fit1_refit$analyses, AIC)),
    mean(sapply(fit2_refit$analyses, AIC)),
    mean(sapply(fit3$analyses, AIC))
  )
)

# --- STAGE 4: Plot the Standardized Comparison ---
ggplot(aic_comparison, aes(x = Model, y = AIC, fill = Model)) +
  geom_bar(stat = "identity", width = 0.5, show.legend = FALSE) +
  geom_text(aes(label = round(AIC, 1)), vjust = -0.5, fontface = "bold") +
  theme_minimal() +
  labs(title = "Model Selection: Standardized AIC Comparison",
       subtitle = paste0("N = ", length(target_patient_ids), " patients (ISPY2/NACT Subset)"),
       y = "Mean AIC",
       x = "") +
  scale_color_manual(values = c("brown1", "royalblue", "darkgreen")) +
  scale_fill_brewer(palette = "Set1")
```

```{r prec-recall, warning=F}
# 1. Helper function to safely extract averaged probabilities from a 'mira' object
get_avg_probs <- function(mira_fit) {
  # Get the number of imputations (m)
  m_count <- length(mira_fit$analyses)
  
  # Extract predictions from each of the m models
  pred_list <- lapply(mira_fit$analyses, function(x) predict(x, type = "response"))
  
  # Average them across rows
  rowMeans(as.data.frame(pred_list))
}

# 2. Calculate probabilities for the two baseline models (Refit on Subset)
avg_probs_m1 <- get_avg_probs(fit1_refit)
avg_probs_m2 <- get_avg_probs(fit2_refit)

# 3. Get the actual pCR values from the subset
actual_pcr <- complete(imputed_collapsed, 1)$pcr

# 4. Create the PR multi-plot object
precrec_obj <- evalmod(
  scores = list(avg_probs_m1, avg_probs_m2, avg_probs_m3), 
  labels = list(actual_pcr, actual_pcr, actual_pcr),
  modnames = c("Model 1: Baseline", "Model 2: Demographic", "Model 3: Treatment")
)

# 5. Plot the Comparison
autoplot(precrec_obj, curvetype = "PRC") +
  theme_minimal() +
  scale_color_manual(values = c("brown1", "royalblue", "darkgreen")) +
  labs(title = "Precision-Recall Comparison: Clinical Impact",
       subtitle = "Evaluation of added predictive value on standardized subset",
       x = "Recall (Sensitivity)",
       y = "Precision (PPV)")
```

```{r forest-plot, warning=F}
# 1. Get the pooled results into a standard dataframe
pooled_df <- as.data.frame(summary(pool(fit3), conf.int = TRUE))

# 2. Rename columns manually if they aren't standard
# Depending on your mice version, column names might be 'estimate' or 'est'
names(pooled_df)[names(pooled_df) == "estimate"] <- "est"
names(pooled_df)[names(pooled_df) == "2.5 %"] <- "low"
names(pooled_df)[names(pooled_df) == "97.5 %"] <- "high"

# 3. Exclude intercept and transform to Odds Ratios
plot_data <- pooled_df[pooled_df$term != "(Intercept)", ]
plot_data$OR <- exp(plot_data$est)
plot_data$OR_low <- exp(plot_data$low)
plot_data$OR_high <- exp(plot_data$high)

# 4. Generate the Forest Plot using standard ggplot2
ggplot(plot_data, aes(x = reorder(term, OR), y = OR)) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "steelblue") +
  geom_pointrange(aes(ymin = OR_low, ymax = OR_high), color = "darkgreen") +
  coord_flip() +
  theme_bw() + # theme_bw is very stable in older ggplot2 versions
  labs(title = "Model 3: Odds Ratios for pCR",
       x = "Predictor", 
       y = "Odds Ratio (95% CI)") +
  scale_y_log10()

```


## Conclusion

...

\newpage
\onecolumn

## References

\vspace{2mm}

::: {#refs}
:::

## Code Appendix

```{r, ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE}
```
